%! Author = partsjoo
%! Date = 16.04.2023


\newpage
\section{Decision}
\textbf{As an election organizer I would choose Q2 using the exponential mechanism.}
The exponential mechanism is more siutable for finding the most popular candidate,
which is the primary goal of the election organizer. For Q2, to achieve a
probability of getting the correct result was for
Wolf with $\epsilon \approx 4.605$ was $\approx 0.99997$ chance (very high probability).
But then again I also guess that's not very secure from a privacy perspective. And our
programmers attempt to add more privacy by changing query Q2 to $\epsilon \approx 0.4605$ might
have been also a bit overreaction.

\textbf{As a voter (interested in privacy) I would choose Q1 using rounding mechanism.}
From a privacy perspective, Q1 using the Laplace mechanism with rounding could be a better choice.
In this scenario, the global sensitivity of the query Q1 is 1, considering there are three candidates.
To ensure the probability of getting the correct result at least 0.9 for each candidate we found that with
$\varepsilon &\approx 1.445$ it was $\approx0.9659$ chance of selecting the right candidate.

\textbf{To conclude}, it is essential to consider the trade-off between privacy and utility in both mechanisms.
With Q2, there might be a higher chance of losing privacy.
On the other hand, with Q1, there might be a higher chance of getting inaccurate results due to rounding.
In either case, it's crucial to weigh the trade-offs between privacy and utility in both scenarios.
