\textbf{HOMEWORK 3: ToC and ToU analysis}\\
\textbf{Mobile app: Strava}\\
\textbf{Online env: Stackoverflow}\\
\textbf{Student:} Joosep Parts 221963IVCM\\

\textbf{stackoverflow.com}

\textbf{What is the minimum age for creating the account in these environments in your home country?} The minimum age for creating an account in the specified online environment varies depending on the location.
In general, you must be at least 13 years old to access or use the services.
However, if you are located within the European Union, the minimum age is raised to 16 years old.
Users under these age limits are not allowed to access or use the services in any manner.

\textbf{How are their services funded?} Advertising: 1) Display ads appear on the website.
Job Listings: Employers pay to post jobs on
the Stack Overflow Jobs
section.
Stack Overflow for Teams: A paid subscription service for private question and answer forums within companies.
Other paid premium services: Stack Overflow also offers paid premium services for job seekers and employers.

\textbf{How many accounts can you legally have and which rules should you follow for the accounts?} They don't explicitly limit the number of accounts a user can have.

\textbf{Who cannot use the services?} The site isn't bias towards any group, ethnicity, religion, etc.
so the statute of limitations is the same for everyone.
However, if you are under the age of 13, you are not allowed to use the services.
It is forbidden to use the services if you are located in a country that is subject to U.S.\ or other applicable export controls and
sanctions.

\textbf{What are the activities against terms of use on this site?} Using multiple accounts for disruptive, abusive, or nefarious behavior would  be against their terms of use.

\textbf{What are the permissions you give to the social media platform?}
Third-party services: The platform collects data through services like Google Ads and Google Analytics for targeted advertising, personalized content, and analytics.

\textbf{Which information about your account will be shared for ads?}
Links and Social Media: If you connect with the platform via social media, both the platform and the social media service may collect additional information like your username, profile picture, and contact list.
Information from other sources: Data is also collected from third parties to combat fraud and provide services.
Contact Forms for Advertisers: If you inquire about advertisements, your name, email, and telephone number are collected.

\textbf{Who are the service providers your information is being shared to for targeted ads?}
Microsoft Advertising, Google Ads, Companies within the Stack Exchange network.

\textbf{Who is allowed to request closing your account in case of your death?}
If Stackoverflow is contacted by a verifiable family member,
they can sometimes make some accommodations on a case-by-case basis depending
on the needs and goals, but they don't offer any guarantee.
This gets especially tricky if the scenario presents internationally.
They have no plans to have a system in place where designated accounts can take certain actions.

\textbf{Strava app}

\textbf{Who is the developer of the app?} The developer of the app Strava is Strava, Inc.

\textbf{What is the developer's business registration code?}
"Strava" of San Francisco, CA 94107 operates primarily in SIC Code 7374 - Computer Processing and Data Preparation and Processing Services and NAICS Code 518210 - Data Processing, Hosting, and Related Services.
Strava is a small-sized business with low revenue, that is well-established in its industry.

\textbf{How many times has the app been downloaded?}
Total Strava mobile app revenue in September 2023 was 10m, including 8m for iOS apps and 2m for Android apps.
Total mobile app downloads for Strava was 2m, including 700k iOS app downloads and 900k Android app downloads.
Strava has a total of 8 apps, including 4 iOS apps, and 4 Android apps.
Their top grossing app last month was Strava: Run and their most downloaded app was Strava: Run.

\textbf{How many people have left a review?}
This information is tricky to find, but according to the Google Play Store, the app has been reviewed by 830k people and 240.8k people have left a review on the Apple App Store.

\textbf{When did the app receive its last update?}
They publish at different rates to different platforms, but the last update to iOS was Version 328.0.0 Oct 11, 2023 Updated on
10/10/2023 for Android.

\textbf{What permissions does the app (may) require and on what purpose (can be longer than 5 sentences)?}
Health Information: To collect or infer health-related data such as heart rate for statistics and visualizations. (access to heart rate
monitor watch, or gyroscope for step counting)
Third-Party Products and Services: To collect and process data from other services or devices like Garmin watch or Peloton, which you connect to the app.
Device-based Settings: To access photos, location, and contacts for the services specified when you enable these settings.
These permissions aim to personalize your experience, provide you with relevant information, and enhance the app's functionalities.
Users can withdraw their consent for these at any time through the app's settings, however it is likely something stops working when you
do it.
For example, I tried to turn off location services but that would completely disable the app's functionality (which is geo-tracking).

\textbf{What is the oddest permission the app requires?}
"Permission to access contacts information through your device-based settings so we can provide the services described when you enable the
settings". What does this even mean!?
I don't want to give access to my contacts to a fitness app when I want to monitor MY health and fitness.


In analyzing the Terms of Conditions and Terms of Use for Stackoverflow and the Strava app, it's evident that both platforms have distinct policies concerning user age, account limitations, and data permissions.
Stackoverflow primarily funds its services through advertising and paid features, with age restrictions that vary by region.
It has a non-biased approach to user limitations and collects data for targeted advertising.

Strava is developed by Strava, Inc. and focuses on health-related functionalities.
It requires a broader set of permissions, including accessing device-based settings like contacts, which odd\ldots

Both platforms offer services that come with a trade-off between functionality and privacy.
While Stackoverflow focuses on professional and educational enrichment, Strava leans more towards health and fitness tracking, requiring more intrusive permissions.
Overall, users should be cautious about the data they're willing to share and should adjust settings accordingly, even if it means losing some functionality.
